### 2020-06-06 model_interventions.Rmd
integrated CHILDES word frequency distribution into model, and now plot learning of some individual words and total vs. CDI words known

---
title: "Modeling Interventions"
output:
  html_document:
    df_print: paged
---

```{r, libraries, include=F}
require(tidyverse)
source("model-nonsampling.R")
```

We'd like to be able to model interventions (both past and proposed) with the standard model. 
How would that work? Let's consider Gilkerson et al. (2017), in which 72 parents of typically developing children aged 9 to 21 months were assigned to immediate- or delayed-treatment (control) conditions. 
During treatment, parents were recorded 10 times over a 3-month period while engaging in a web-based intervention offering strategies for increasing talk and interaction. 
During a 9-month follow-up, parents completed additional recordings and language assessments. 
Overall, no differences were found in language behaviors between immediate-treatment versus delayed-treatment groups. 
However, parents who were at first below-average on automated language measures showed significant postintervention increases which held longitudinally, and children showed significant elevations in language ability.
Problem: Gilkerson 2017 only reports standardized scores (mean = 100 and SD = 15) for AWC (median split: bottom half M=84.09, upper half M=116.01), which are calculated from the original LENA norms (Gilkerson & Richards, 2008), where they report overall M=12,297 (SD=6462) per 12-hour session (so 1025 words/hour).

```{r, summary-tables, echo=F}
# Gilkerson et al. 2017 Table 2: standardized AWC

# Gilkerson et al. 2018 Table 2: 
tribble(
  ~age_mos, ~AWC_mean, ~AWC_sd,
  "2-17",   14484, 6022,
  "18-24",  11768, 4159,
  "25+",   12928, 4342
) # total sample (N=146): 13218, sd=5079, range 3372-30794

# should look at Ramírez-Esparza, García-Sierra, and Kuhl (2014, 2017)
# ADS / CDS 30s LENA segments

# on average, Bergelson2018 (North American) children heard
# 11.36 min of CDS per hour (SD = 4.24)
# average of 7.3 min of ADS per hour (SD = 6.4)
# on average 65% of the speech infants heard from adults sounded child directed (SD = 0.22); varied widely, with one infant hearing only 17% CDS while another heard 100% CDS
# CDS quantity remained largely stable with child age, ADS quantity decreased with child age
# "Shneidman et al. (2013) finds 69% CDS in North American homes of 14- to 42 month olds while we find 65% on average across 3–20 months"

# Bergelson et al. 2018 Table S1: 
Bergelson_s1 <- tribble(
  ~study, ~AWC_mean, ~AWC_sd, ~CDS_mean, ~CDS_sd, ~N, 
  "Schneidman & Goldin-Meadow 2013", 6254, NA, 2063, NA, 15,
  "Schneidman & Goldin-Meadow 2013", NA, NA, 2404, NA, 15,
  "Weisleder & Fernald 2013", 1084, 563, 598, 367, 29,
  "Johnson et al. 2014", 1725, NA, NA, NA, 33, # also a group of 7 with AWC ~1000/hr
  "Tamis-LeMonda et al. 2017", 2197, NA, NA, NA, 40,
  "Pancsofar & Vernon-Feagans (2006)", 2239, NA, NA, NA, 92, # averaged together mom (2559) and dad (1919)
  "Gilkerson et al. 2017", 1059, 356, NA, NA, 329,
  "Hart & Risley", NA, NA, NA, NA, NA
)

# Table S2: prev studies' LENA results, along with the 59 of 61 recordings in the current dataset that are >8hrs
Bergelson_s2 <- tribble(
  ~study, ~AWC_mean, ~AWC_sd, ~CVC_mean, ~CVC_sd, ~N,
  "Gilkerson et al. 2017", 12709, 4274, 1817, 787, 72, # these are full-day recordings (12hrs)
  "Greenwood et al. 2011", 13142, 5562, 1714, 850, 30, # mid- to upper-SES; AWC range: 631-36563, CVC range: 11-5611
  "Soderstrom & Wittebolle", 10125, 4890, 1744, 1058, 11, # mostly shorter recordings (6-8hrs)
  "Zimmerman et al. 2009", 12800, 4400, NA, NA, 275, # cross-sectional; only 71 longitudinal
  "Bergelson et al. 2018", 16510, 8718, 1432, 764, 59,
)
# from Melanie and Elika and Alex Cristia:
# N children = 264; N recordings = 851 AWC/hr across the recordings/participants:
#    0%    25%    50%      75%     100%
#  33.63  756    1087     1458    3440

Bergelson_s2 <- Bergelson_s2 %>% mutate(AWChr_mean = AWC_mean / 12,
                                        AWChr_sd = AWC_sd / 12)

# Sperry et al. Figure 1: HR = Hart & Risley 1995
# words/hour spoken by primary caregivers to children
Sperry <- tribble(
  ~group, ~CDS_mean, ~N, ~AllCaregiverPctIncr, ~AmbientEnvPctIncr, #AllCaregiver_mean, AmbientEnv_mean,
  "HR: Kansas professionals", 2153, 13, NA, NA, #NA, NA,
  "Black Belt Poor", 1838, 11, 1.58, 2.02, # NA, 3203,
  "Longwood Middle Class", 1491, 6, 1.30, 2.04, # 1777, NA
  "HR: Kansas Middle Class", 1400, 10, NA, NA,
  "Daly Park Working Class", 1351, 7, 1.17, 1.75,
  "HR: Kansas Lower Class", 1137, 13, NA, NA,
  "South Baltimore Poor", 1061, 3, 1.21, 1.54,
  "Jefferson Working Class", 1048, 15, 1.53, 3.10,
  "Kansas Welfare", 616, 6, NA, NA
)
# final columns from Table 2 let us recreate Figure 2, which breaks down into Primary Caregiver / All Caregivers / Ambient Environment
# e.g., Black Belt Poor Ambient Env = 1838 * 2.02 should be 3203 words/hr (as reported in text..) but is 3713..

Sperry <- Sperry %>% mutate(AllCaregiver = CDS_mean * AllCaregiverPctIncr,
                  AmbientEnv = CDS_mean * AmbientEnvPctIncr)
# "Finally, children in the poor Black Belt community heard an astounding 3,203 words per hour in their ambient environment, a 102% increase over the number of words addressed to them by their primary caregivers."

print(Bergelson_s2)
```

Let's use the hourly rates from individual children from Sperry et al. (2018), Hart & Risley (1995), Soderstrom & Wittebolle (2013), and Weisleder & Fernald (2013) to estimate the parameters of hourly input rate distributions for both child-directed speech (CDS) and all adult speech.

```{r, cds, echo=F, warning=F, caption="Distribution of hourly child-directed speech rates heard by individual children."}
df <- read.csv("data/hourly_tokens_Sperry_HartRisley.csv")
cds_agg <- df %>% summarise(mean = mean(mother_child_tokens_hr, na.rm=T),
                           sd = sd(df$mother_child_tokens_hr, na.rm=T))
# CDS mean: 1197.5, sd: 840.2
breaks = seq(0, 3650, by = 120) 
ggplot(df, aes(x = mother_child_tokens_hr)) + 
    geom_histogram(aes(y =..density..), breaks = breaks, 
                   color = "black", fill = "gray") +
    stat_function(fun = dnorm, args = list(mean = cds_agg$mean, sd = cds_agg$sd)) +
  geom_vline(xintercept=cds_agg$mean, lty="dashed") + xlab("Child-directed Speech (Tokens/Hour)") + theme_bw() +
  annotate("text", x=cds_agg$mean +700, y=.0008, label=paste("Mean =",round(cds_agg$mean,0)))
ggsave("figures/CDS_hourly_tokens_Sperry_HartRisley_WeislederFernald.pdf", width=4.5,height=4)

# should this be a Poisson distribution? not normal
#ggplot(transform(data.frame(x=0:3650), y=dpois(x, cds_agg$mean)), aes(x, y)) + 
#    geom_bar(stat="identity")
```

```{r, awc, echo=F, warning=F, caption="Distribution of hourly adult speech rates heard by individual children."}
# all (adult) tokens Sperry + Soderstrom Wittebolle + Weisleder Fernald
awc_agg <- df %>% summarise(mean = mean(all_tokens_hr, na.rm=T),
                           sd = sd(df$all_tokens_hr, na.rm=T))
# AWC: mean: 1797.8 sd=1130.2
breaks = seq(0, 5450, by = 150) 
ggplot(df, aes(x = all_tokens_hr)) + 
    geom_histogram(aes(y =..density..), breaks = breaks, 
                   color = "black", fill = "gray") + theme_bw() +
    stat_function(fun = dnorm, args = list(mean = awc_agg$mean, sd = awc_agg$sd)) +
  geom_vline(xintercept=awc_agg$mean, lty="dashed") + xlab("Adult Word Count (Tokens/Hour)") +
  annotate("text", x=awc_agg$mean +700, y=.0008, label=paste("Mean =",round(awc_agg$mean,0)))
ggsave("figures/all_hourly_tokens_Sperry_SoderstromWittebolle_WeislederFernald.pdf", width=5,height=4)

nsubj_cds = sum(!is.na(df$mother_child_tokens_hr)) # 113 children
nsubj_awc = sum(!is.na(df$all_tokens_hr)) # 83 children
```

Based on data from `r nsubj_cds` children, the mean number of child-directed tokens heard per hour is `r round(cds_agg$mean,0)` (sd=`r round(cds_agg$sd,0)`).
Using data from `r nsubj_awc` children, the mean number of all adult tokens heard per hour is `r round(awc_agg$mean,0)` (sd=`r round(awc_agg$sd,0)`).

## References

Gilkerson, J., Richards, J. A., & Topping, K. (2017). Evaluation of a lena-based online intervention for parents of young children. Journal of Early Intervention, 1053815117718490.

Greenwood, C. R., Thiemann-Bourque, K., Walker, D., Buzhardt, J., & Gilkerson, J. (2011). Assessing children’s home language environments using automatic speech recognition technology. Communication Disorders Quarterly, 32(2), 83-92.

Soderstrom, M., & Wittebolle, K. (2013). When do caregivers talk? the influences of activity and time of day on caregiver speech and child vocalizations in two childcare environments. PloS one, 8(11), e80646.

Weisleder, A., & Fernald, A. (2013). Talking to Children Matters: Early Language Experience Strengthens Processing and Builds Vocabulary. Psychological Science, 24(11), 2143–2152.

Zimmerman, F. J., Gilkerson, J., Richards, J. A., Christakis, D. A., Xu, D., Gray, S., & Yapanel, U. (2009). Teaching by listening: The importance of adult-child conversations to language development. Pediatrics, 124(1), 342–349.