### 2021-09-24 childrens-input.Rmd
estimating total language input

---
title: "Children's language input"
author: "George"
date: "9/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

# Goal

Use average hourly input rates from three sources (overheard speech, child-directed speech, and child-read books) to estimate how many tokens a 'standard' child will experience over their early development.

## Overheard and Child-directed speech

Using children's median number of waking hours, we extrapolate from our best estimates of hourly rates for overheard speech (600 words/hr) and child-directed speech (1198 words/hr) to the monthly amounts of speech heard by children.
(Should these rates change for school-aged children?)

```{r waking-hours}
# https://www.sleepfoundation.org/children-and-sleep/how-much-sleep-do-kids-need
waking_hours_per_day = c(7,8,9, # 14-17 hrs of sleep
                           seq(10, 12, length.out=9), # infant: 12-15 hrs of sleep
                           rep(12, 12), # toddler: 11-14 hrs of sleep
                           rep(13, 3*12), # preschool (3-5): 10-13 hrs of sleep
                           rep(14, 8*12)) # school-aged: 9-11 hrs of sleep
```

```{r hourly-rates}
overheard_speech_rate = rep(600, 13*12) # 13 years - increase at school age?
child_directed_speech_rate = rep(1198, 13*12) # how should this change?
# let's estimate monthly rates based on Hasbrauck and Tindal (2006) norms
reading_rate = 60 * c(rep(0, 12*5), # no reading 0-5
                 seq(0, 53, length.out=12), # 53 correct wpm by end of Grade 1
                 seq(53, 89, length.out=12), # 89 by end of Grade 2
                 seq(89, 107, length.out=12), # 107 by end of Grade 3
                 seq(107, 123, length.out=12), # Grade 4
                 seq(123, 139, length.out=12), # Grade 5
                 seq(139, 150, length.out=12), # Grade 6
                 seq(150, 150, length.out=12), # Grade 7
                 seq(150, 151, length.out=12) # Grade 8
                 ) # first 13 years

# 6TH-8TH GRADE (SPRING) (11-14 YEARS OLD): 150 – 204 WPM
# HIGHSCHOOL (14-18 YEARS OLD): 200 – 300 WPM
# COLLEGE (18-23) YEARS OLD	300 – 350 WPM
# ADULTS	220 – 350 WPM
```

## Reading rates

Many sources recommend that school-aged children read 15-20 minutes per day (7.6 hours/month), although most teachers estimate that most of their students read less than this.
For now we will assume the 'standard' child reads the recommended daily amount, at speeds following the Hasbrauck and Tindal (2006) norms (e.g., 53 correct words/min by end of Grade 1; 151 correct words/min by end of Grade 8).
For reference, silent-reading adults average 238 words/min, while adults that read aloud average 183 words/min (Brysbaert, 2019).

```{r, echo=FALSE}
hours_read_per_month = 30.44*.25 # 15 min/day
tokens_read_per_month = reading_rate * hours_read_per_month
waking_hours_per_month = waking_hours_per_day * 30.44

hours_overheard_speech_per_month = overheard_speech_rate * waking_hours_per_month
hours_child_directed_speech_per_month = child_directed_speech_rate * waking_hours_per_month

input_long_df <- bind_rows(
  tibble(Source = "Reading", Age = 1:length(tokens_read_per_month), 
         `Hourly Rate` = reading_rate,
         `Monthly Tokens` = tokens_read_per_month,
         `Total Tokens` = cumsum(tokens_read_per_month)),
  tibble(Source = "Overheard Speech", Age = 1:length(tokens_read_per_month),
         `Hourly Rate` = overheard_speech_rate,
         `Monthly Tokens` = hours_overheard_speech_per_month,
         `Total Tokens` = cumsum(hours_overheard_speech_per_month)),
  tibble(Source = "Child-directed Speech", Age = 1:length(tokens_read_per_month),
         `Hourly Rate` = child_directed_speech_rate,
         `Monthly Tokens` = hours_child_directed_speech_per_month,
         `Total Tokens` = cumsum(hours_child_directed_speech_per_month))
  )
```

## Monthly input

For reference, the average novel is ~60,000 words.

```{r monthly-input, echo=F}
input_long_df %>%
  mutate(`Age (years)` = Age / 12) %>%
  ggplot(aes(x=`Age (years)`, y=`Monthly Tokens`, fill=Source)) + 
    geom_area() + #ylim(0, 1.5e6) + 
    theme_classic() # + geom_hline(yintercept=1e6, linetype="dashed")
```

## Cumulative input

How many tokens do children accumulate from different sources by the age of 13?

```{r, echo=F}
input_long_df %>% 
  mutate(`Cumulative Tokens (millions)` = `Total Tokens` / 1e6,
         `Age (years)` = Age / 12) %>%
  ggplot(aes(x=`Age (years)`, y=`Cumulative Tokens (millions)`, fill=Source)) + 
    geom_area() + 
    theme_classic() + # + geom_hline(yintercept=1e6, linetype="dashed")
    theme(legend.position = c(.2, .7))
```

(Add TV? "The average across all 170 stations is 2.38 words per second" (143 words/min, 8580 words/hr). The average child (aged 8-12) watches 4-6 hours/day -- 34320 to 51480 words that are not overheard/child-directed speech.)